---
title: "Untitled"
author: "Caitie"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(ggplot2)
library(here)
library(tidyverse)
library(terra)
library(cowplot)
```


#Colours 

```{r}
colPalette <- c("#3B3561", "#ECA72C", "#226F54","#B8D4E3", "#C06E52","#01172F", "#006D77")

``` 


# Plot PE through time

```{r}
pe<-read_csv(here("output_data/results/PE_baseline_all_imcra.csv"))
```

##Create new data frame for PE
```{r}
pe_long <- pivot_longer(pe, cols= c("CAPAD_PE", "Strict_PE", "Non_Strict_PE"),
                         names_to = "PE_type",
                         values_to = "PE")

pe_long<- pe_long%>% 
  dplyr::select(-c(CAPAD_N,Strict_N,Non_Strict_N))
```

##Plot
```{r}

ggplot(pe_long, aes(x = Year, y = PE, col = PE_type)) +
  geom_line() +
  theme_bw() +
  scale_colour_discrete(type = colPalette, labels = c("All MPAs", "Non-strict", "Strict")) +
  labs(title = "Ecoregion Protection Equality - Baseline",
      y = "Protection equality",
      x = "Year", 
      col = "Protected area class")
   

```



```{r}
ggsave(here("figures/All_PAs_Protection_equality.png"), width = 10, height= 6)
```

# Plot corrected PE through time 

```{r}
pe_corrected<-read_csv(here("output_data/results/PE_corrected_all_imcra.csv"))
```

##Create new data frame for PE
```{r}
pe_corrected_long <- pivot_longer(pe_corrected, cols= c("Capad_PE", "Strict_PE", "Non_Strict_PE"),
                         names_to = "PE_type",
                         values_to = "PE")

pe_corrected_long<- pe_corrected_long%>% 
  dplyr::select(-c(Capad_N,Strict_N,Non_Strict_N))
```


##Plot
```{r}

ggplot(pe_corrected_long, aes(x = Year, y = PE, col = PE_type)) +
  geom_line() +
  theme_bw() +
  scale_colour_discrete(type = colPalette, labels = c("All MPAs", "Non-strict", "Strict")) +
  labs(title = "Ecoregion Protection Equality - Corrected",
      y = "Protection equality",
      x = "Year", 
      col = "Protected area class")
   

```


```{r}
ggsave(here("figures/Corrected_Protection_equality.png"), width = 10, height= 6)
```


#Plot by protection type 
##Strict
```{r}

pe_long_strict <- pe_long%>%
  group_by(PE_type)%>%
  filter(PE_type == "Strict_PE")

pe_corrected_long_strict <- pe_corrected_long%>%
  group_by(PE_type)%>%
  filter(PE_type == "Strict_PE")

```


```{r}
ggplot() +
  theme_bw() +
  geom_line(data = pe_long_strict, aes(x = Year, y = PE, col = "Baseline")) +
  geom_line(data = pe_corrected_long_strict, aes(x = Year, y = PE, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(title = "Strict",
      y = "Protection equality",
      x = "Year",
      col = "Type")

```

```{r}
ggsave(here("figures/Strict_PE_all.png"), width = 10, height= 6)
```

##Non strict
```{r}

pe_long_non_strict <- pe_long%>%
  group_by(PE_type)%>%
  filter(PE_type == "Non_Strict_PE")

pe_corrected_long_non_strict <- pe_corrected_long%>%
  group_by(PE_type)%>%
  filter(PE_type == "Non_Strict_PE")

```


```{r}
ggplot() +
  theme_bw() +
  geom_line(data = pe_long_non_strict, aes(x = Year, y = PE, col = "Baseline")) +
  geom_line(data = pe_corrected_long_non_strict, aes(x = Year, y = PE, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(title = "Non-strict",
      y = "Protection equality",
      x = "Year",
      col = "Type")

```

```{r}
ggsave(here("figures/Non_strict_PE_all.png"), width = 10, height= 6)
```

##Full CAPAD network
```{r}

pe_long_capad <- pe_long%>%
  group_by(PE_type)%>%
  filter(PE_type == "CAPAD_PE")

pe_corrected_long_capad <- pe_corrected_long%>%
  group_by(PE_type)%>%
  filter(PE_type == "Capad_PE")

```


```{r}
ggplot() +
  theme_bw() +
  geom_line(data = pe_long_capad, aes(x = Year, y = PE, col = "Baseline")) +
  geom_line(data = pe_corrected_long_capad, aes(x = Year, y = PE, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(title = "Full CAPAD network",
      y = "Protection equality",
      x = "Year",
      col = "Type")

```


```{r}
ggsave(here("figures/Capad_PE_all.png"), width = 10, height= 6)
```

#Plot MPG 
##Load in data 

```{r}
mpg<- read_csv(here("output_data/results/MPG_baseline_all_imcra.csv"))
```

##Create new data frame for MPG 
```{r}

mpg_long <- pivot_longer(mpg, cols= c("Capad_MPG", "Strict_MPG", "Non_Strict_MPG"),
                         names_to = "MPG_type",
                         values_to = "MPG")
mpg_long<- mpg_long%>% 
  dplyr::select(-c(Capad_N,Strict_N,Non_Strict_N))

```

##Plot
```{r}

ggplot(mpg_long, aes(x = Year, y = MPG, col = MPG_type)) +
  geom_line() +
  theme_bw() +
  scale_colour_discrete(type = colPalette, labels = c("All MPAs", "Non-strict", "Strict")) +
  labs(title = "Mean Protection Gap - Baseline",
      y = "Mean Protection Gap",
      x = "Year",
      col = "Protected area class")
   
```

```{r}
ggsave(here("figures/All_PAs_MPG.png"), width =10, height=6)
```



#Plot corrected MPG
```{r}
mpg_corrected<- read_csv(here("output_data/results/MPG_corrected_all_imcra.csv"))
```

##Create new data frame for MPG 
```{r}

mpg_corrected_long <- pivot_longer(mpg_corrected, cols= c("Capad_MPG", "Strict_MPG", "Non_Strict_MPG"),
                         names_to = "MPG_type",
                         values_to = "MPG")
mpg_corrected_long<- mpg_corrected_long%>% 
  dplyr::select(-c(Capad_N,Strict_N,Non_Strict_N))

```

##Plot
```{r}

ggplot(mpg_corrected_long, aes(x = Year, y = MPG, col = MPG_type)) +
  geom_line() +
  theme_bw() +
  scale_colour_discrete(type = colPalette, labels = c("All MPAs", "Non-strict", "Strict")) +
  labs(title = "Mean Protection Gap - Corrected",
      y = "Mean Protection Gap",
      x = "Year",
      col = "Protected area class")
   
```

```{r}
ggsave(here("figures/Corrected_MPG.png"), width =10, height=6)
```


#Plot by protection type 
##Strict
```{r}
mpg_long_strict <- mpg_long%>%
  group_by(MPG_type)%>%
  filter(MPG_type == "Strict_MPG")

mpg_corrected_long_strict <- mpg_corrected_long%>%
  group_by(MPG_type)%>%
  filter(MPG_type == "Strict_MPG")

```


```{r}
ggplot() +
  theme_bw() +
  geom_line(data = mpg_long_strict, aes(x = Year, y = MPG, col = "Baseline")) +
  geom_line(data = mpg_corrected_long_strict, aes(x = Year, y = MPG, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(title = "Strict",
      y = "Mean protection gap",
      x = "Year",
      col = "Type")

```


```{r}
ggsave(here("figures/Strict_MPG_all.png"), width =10, height=6)
```


##Non strict
```{r}
mpg_long_non_strict <- mpg_long%>%
  group_by(MPG_type)%>%
  filter(MPG_type == "Non_Strict_MPG")

mpg_corrected_long_non_strict <- mpg_corrected_long%>%
  group_by(MPG_type)%>%
  filter(MPG_type == "Non_Strict_MPG")

```


```{r}
ggplot() +
  theme_bw() +
  geom_line(data = mpg_long_non_strict, aes(x = Year, y = MPG, col = "Baseline")) +
  geom_line(data = mpg_corrected_long_non_strict, aes(x = Year, y = MPG, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(title = "Non-strict",
      y = "Mean protection gap",
      x = "Year",
      col = "Type")

```


```{r}
ggsave(here("figures/Non_strict_MPG_all.png"), width =10, height=6)
```


##Full Capad Network 
```{r}
mpg_long_capad <- mpg_long%>%
  group_by(MPG_type)%>%
  filter(MPG_type == "Capad_MPG")

mpg_corrected_long_capad <- mpg_corrected_long%>%
  group_by(MPG_type)%>%
  filter(MPG_type == "Capad_MPG")

```


```{r}
ggplot() +
  theme_bw() +
  geom_line(data = mpg_long_capad, aes(x = Year, y = MPG, col = "Baseline")) +
  geom_line(data = mpg_corrected_long_capad, aes(x = Year, y = MPG, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(title = "Full Capad Network",
      y = "Mean protection gap",
      x = "Year",
      col = "Type")

```


```{r}
ggsave(here("figures/Capad_MPG_all.png"), width = 10, height= 6)
```


#Total area 
##Read in area data 
```{r}
baseline_capad_area <- read_csv(here("output_data/Total_area/Baseline_capad_area.csv"))

baseline_strict_area <- read_csv(here("output_data/Total_area/Baseline_strict_area.csv"))

baseline_non_strict_area <- read_csv(here("output_data/Total_area/Baseline_non_strict_area.csv"))

```

```{r}
adjusted_capad_area <- read_csv(here("output_data/Total_area/Adjusted_capad_area.csv"))

adjusted_strict_area <- read_csv(here("output_data/Total_area/Adjusted_strict_area.csv"))

adjusted_non_strict_area <- read_csv(here("output_data/Total_area/Adjusted_non_strict_area.csv"))


```

##Create new data frame for area 
```{r}

baseline_capad <- baseline_capad_area%>% 
  dplyr::select(-c(...1, GIS_AREA, area_km2))

baseline_strict<- baseline_strict_area%>% 
  dplyr::select(-c(...1, GIS_AREA, area_km2))

baseline_non_strict<- baseline_non_strict_area%>% 
  dplyr::select(-c(...1, GIS_AREA, area_km2))

```

##Capad area figures
```{r}
Area1<- ggplot() +
  theme_bw() +
  geom_line(data = baseline_capad, aes(x = Year, y = total_area_km2, col = "Baseline")) +
  geom_line(data = adjusted_capad_area, aes(x = Year, y = Capad_area, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
    theme(plot.margin = margin(r = 10))+
  labs(title = "Full Capad Network",
      y = "Area (km2)",
      x = "Year",
      col = "Type")

```


```{r}
ggsave(here("figures/Capad_area.png"), width = 10, height= 6)
```


```{r}
Area2<- ggplot() +
  theme_bw() +
  geom_line(data = baseline_strict, aes(x = Year, y = total_area_km2, col = "Baseline")) +
  geom_line(data = adjusted_strict_area, aes(x = Year, y = Strict_area, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
    theme(plot.margin = margin(r = 10))+
  labs(title = "Strict",
      y = "Area (km2)",
      x = "Year",
      col = "Type")

```


```{r}
ggsave(here("figures/Strict_area.png"), width = 10, height = 6)
```

```{r}
Area3<- ggplot() +
  theme_bw() +
  geom_line(data = baseline_non_strict, aes(x = Year, y = total_area_km2, col = "Baseline")) +
  geom_line(data = adjusted_non_strict_area, aes(x = Year, y = Non_strict_area, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
    theme(plot.margin = margin(r = 10))+
  labs(title = "Non-Strict",
      y = "Area (km2)",
      x = "Year",
      col = "Type")

```

```{r}
Arealegend<- ggplot() +
  theme_bw() +
  geom_line(data = baseline_non_strict, aes(x = Year, y = total_area_km2, col = "Baseline")) +
  geom_line(data = adjusted_non_strict_area, aes(x = Year, y = Non_strict_area, col = "Corrected")) +
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(title = "Non-Strict",
      y = "Area (km2)",
      x = "Year",
      col = "Type")

Legend<- get_legend(Arealegend)
```


```{r}
ggsave(here("figures/Non_strict_area.png"), width = 10, height = 6)
```


```{r}
plot_grid(Area1, Area2, Area3, Legend)

plot_grid(plot_grid(PE1, MPG1, PE2, MPG2, PE3, MPG3, labels = c("a) Full Capad network","", "b) Non-strict","","c) Strict"), label_size = 12, ncol=2, hjust = 0), legend, rel_widths = c(1, 0.2))
```


#Metrics per area 
## merge dataframes 

```{r}
eez<- vect(here("output_data/projected_eez.shp"))

eez_area<- sum(expanse(eez, unit = "km"))
```



##PE capad
```{r}
baseline_merge1<- merge(baseline_capad, pe_long_capad, by = "Year", all = "TRUE")
baseline_merge_capad<- merge(baseline_merge1, mpg_long_capad, by = "Year", all = "TRUE")

baseline_capad_mutate<- baseline_merge_capad%>%
  mutate(Prop = total_area_km2/eez_area)
 

adjust_merge<- merge(adjusted_capad_area, pe_corrected_long_capad, by = "Year", all = "TRUE") 
adjust_merge_capad<- merge(adjust_merge, mpg_corrected_long_capad,by = "Year", all = "TRUE" )

adjust_capad_mutate<- adjust_merge_capad%>%
  mutate(Prop = Capad_area/eez_area)
```


```{r}
PE1<- ggplot() +
  theme_bw() +
  geom_point(data = baseline_capad_mutate, aes(x = Prop, y = PE, col = "Baseline"), na.rm = TRUE)+
  geom_smooth(data = baseline_capad_mutate, aes(x = Prop, y = PE, col = "Baseline"), na.rm = TRUE, method = "lm", se = FALSE)+
  geom_point(data = adjust_capad_mutate, aes(x = Prop, y = PE, col = "Corrected"), na.rm = TRUE)+
geom_smooth(data = adjust_capad_mutate, aes(x = Prop, y = PE, col = "Corrected"), na.rm = TRUE, method = "lm", se = FALSE)+
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
    theme(plot.margin = margin(r = 10))+
  labs(title = "",
      y = "PE",
      x = "Proportion",
      col = "Type")
```

##PE non_strict
```{r}
baseline_merge1<- merge(baseline_non_strict, pe_long_non_strict, by = "Year", all = "TRUE")
baseline_merge_non_strict<- merge(baseline_merge1, mpg_long_non_strict, by = "Year", all = "TRUE")

baseline_non_strict_mutate<- baseline_merge_non_strict%>%
  mutate(Prop = total_area_km2/eez_area)
 

adjust_merge<- merge(adjusted_non_strict_area, pe_corrected_long_non_strict, by = "Year", all = "TRUE") 
adjust_merge_non_strict<- merge(adjust_merge, mpg_corrected_long_non_strict,by = "Year", all = "TRUE" )

adjust_non_strict_mutate<- adjust_merge_non_strict%>%
  mutate(Prop = Non_strict_area/eez_area)
```


```{r}
PE2<- ggplot() +
  theme_bw() +
  geom_point(data = baseline_non_strict_mutate, aes(x = Prop, y = PE, col = "Baseline"), na.rm = TRUE)+
  geom_smooth(data = baseline_non_strict_mutate, aes(x = Prop, y = PE, col = "Baseline"), na.rm = TRUE, method = "lm", se = FALSE)+
  geom_point(data = adjust_non_strict_mutate, aes(x = Prop, y = PE, col = "Corrected"), na.rm = TRUE)+
geom_smooth(data = adjust_non_strict_mutate, aes(x = Prop, y = PE, col = "Corrected"), na.rm = TRUE, method = "lm", se = FALSE)+
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
    theme(plot.margin = margin(r = 10))+
  labs(title = "",
      y = "PE",
      x = "Proportion",
      col = "Type")
```

##PE strict
```{r}
baseline_merge1<- merge(baseline_strict, pe_long_strict, by = "Year", all = "TRUE")
baseline_merge_strict<- merge(baseline_merge1, mpg_long_strict, by = "Year", all = "TRUE")

baseline_strict_mutate<- baseline_merge_strict%>%
  mutate(Prop = total_area_km2/eez_area)
 

adjust_merge<- merge(adjusted_strict_area, pe_corrected_long_strict, by = "Year", all = "TRUE") 
adjust_merge_strict<- merge(adjust_merge, mpg_corrected_long_strict,by = "Year", all = "TRUE" )

adjust_strict_mutate<- adjust_merge_strict%>%
  mutate(Prop = Strict_area/eez_area)
```


```{r}
PE3<- ggplot() +
  theme_bw() +
  geom_point(data = baseline_strict_mutate, aes(x = Prop, y = PE, col = "Baseline"), na.rm = TRUE)+
  geom_smooth(data = baseline_strict_mutate, aes(x = Prop, y = PE, col = "Baseline"), na.rm = TRUE, method = "lm", se = FALSE)+
  geom_point(data = adjust_strict_mutate, aes(x = Prop, y = PE, col = "Corrected"), na.rm = TRUE)+
geom_smooth(data = adjust_strict_mutate, aes(x = Prop, y = PE, col = "Corrected"), na.rm = TRUE, method = "lm", se = FALSE)+
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
    theme(plot.margin = margin(r = 10))+
  theme(legend.position = "none")+
  labs(title = "",
      y = "PE",
      x = "Proportion",
      col = "Type")
```


##MPG capad
```{r}
MPG1<- ggplot() +
  theme_bw() +
  geom_point(data = baseline_capad_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE)+
  geom_smooth(data = baseline_capad_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE, method = "lm", se = FALSE)+
  geom_point(data = adjust_capad_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE)+
geom_smooth(data = adjust_capad_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE, method = "lm", se = FALSE)+
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
    theme(plot.margin = margin(r = 10))+
  labs(title = "",
      y = "MPG",
      x = "Proportion",
      col = "Type")
```


##MPG non_strict 
```{r}
MPG2<- ggplot() +
  theme_bw() +
  geom_point(data = baseline_non_strict_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE)+
  geom_smooth(data = baseline_non_strict_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE, method = "lm", se = FALSE)+
  geom_point(data = adjust_non_strict_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE)+
geom_smooth(data = adjust_non_strict_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE, method = "lm", se = FALSE)+
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
    theme(plot.margin = margin(r = 10))+
  labs(title = "",
      y = "MPG",
      x = "Proportion",
      col = "Type")
```


##MPG strict

```{r}
MPG3<- ggplot() +
  theme_bw() +
  geom_point(data = baseline_strict_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE)+
geom_smooth(data = baseline_strict_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE, method = "lm", se = FALSE)+
  geom_point(data = adjust_strict_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE)+
geom_smooth(data = adjust_strict_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE, method = "lm", se = FALSE)+
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  theme(legend.position = "none")+
  theme(plot.margin = margin(r = 10))+
  labs(title = "",
      y = "MPG",
      x = "Proportion",
      col = "Type")
MPG3

```


```{r}
MPGlegend<- ggplot() +
  theme_bw() +
  geom_point(data = baseline_strict_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE)+
  geom_smooth(data = baseline_strict_mutate, aes(x = Prop, y = MPG, col = "Baseline"), na.rm = TRUE, method = "lm", se = FALSE)+
  geom_point(data = adjust_strict_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE)+
geom_smooth(data = adjust_strict_mutate, aes(x = Prop, y = MPG, col = "Corrected"), na.rm = TRUE, method = "lm", se = FALSE)+
  scale_colour_discrete(type = colPalette, labels = c("Baseline", "Corrected")) + 
  labs(y = "MPG",
      x = "Proportion",
      col = "Type")

legend<- get_legend(MPGlegend)
```



```{r}
plot_grid(plot_grid(PE1, MPG1, PE2, MPG2, PE3, MPG3, labels = c("a) Full Capad network","", "b) Non-strict","","c) Strict"), label_size = 12, ncol=2, hjust = 0), legend, rel_widths = c(1, 0.2))

ggsave(here("figures/MetricvsProp.png"))

```

