---
title: "Untitled"
author: "Caitie"
date: "2024-07-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(terra)
library(sf)
library(here)
library(pbmcapply)
library(parallel)
library(foreach)
library(doParallel)
```

# Load datasets

After looking at the terrestrial footprint data, I think we should focus on "direct" pressures. The marine footprint is a bit different as it includes climate change threats, which MPAs cannot abate. Since we are focused on how effective MPAs are - I think it makes sense to only focus on threats they can stop (so direct threats that MPAs can stop).

These include (based on Jones et al. wilderness paper)

All forms of fishing (6 in total)
Benthic structures # Excluded in the pace of change paper due to data limitations
Commercial Shipping
Invasive Species #Excluded in the pace of change paper due to data limitations
Ocean-based pollution # Excluded in the pace of change paper due to data limitations

We can also exclude land-based pressures because MPAs cannot stop them either so that leaves us with 7 pressures.


We then want to (as per Jones et al.) for each year is:

1. Identify the lowest 10% of each layer (<10% = 0, >=10% = 1)

We want to use the normalised (or rescaled) layer for each pressure that is between 0 and 1.

```{r}
years<-c(2003:2013)

stressors<-list.files(path = here("raw_data/HFP/stressors"),full.name = T)

```

```{r}
registerDoParallel(5)
foreach(stress = 1:length(stressors)) %dopar%{ # year=2013 for some reason this was not running automatically and I had to run manually in chunks.

   t<-rast(stressors[[stress]])
    name<-names(t)
    
    t2<-t
    t2[t2==0]<-NA #we may need to think about this, since there are so many zeros in some layers
    
    c_10<-global(t2, \(i) quantile(i, 0.1, na.rm=T))

    c_corr<-ifel(t2<c_10[[1]], 0, 1, filename = here(paste0("output_data/HFP/wilderness/stressors/Wilderness_", name,".tif")),overwrite = T) # 0 is wilderness and 1 is not
    
    #writeRaster(c_corr, here(paste0("output_data/HFP/wilderness/stressors/Wilderness_", name,".tif")),overwrite = T)
  
}

```



2. Exclude the lowest 10% of cumulative impact in that year

I think we will need to recalculate the impact layer based on only the direct threats

```{r}
chi_list<-list.files(path = here("output_data/HFP/cumulative_impact"),full.name = T)

```

```{r}
registerDoParallel(5)
foreach(chi = 1:length(chi_list)) %dopar%{ 
chi_sub<-rast(chi_list[[chi]])

name<-names(chi_sub)

chi_sub2<- chi_sub
chi_sub2[chi_sub2==0]<-NA 

chi_10_2<-global(chi_sub2, \(i) quantile(i, 0.1, na.rm=T))

chi_corr<-ifel(chi_sub2<chi_10_2[[1]], 0, 1, filename = here(paste0("output_data/HFP/wilderness/direct_chi/Wildnerness_",name,".tif")), overwrite = TRUE) # wilderness is now 0 and everything else is 1


}
```

3. Sum layers to identify areas that are in the lowest 10% for each layer (so value of 0 after being summed)

```{r}
years<-c(2003:2013)
wild_chi<-list.files(here("output_data/HFP/wilderness/direct_chi"), full = T)
wild_stress<-list.files(here("output_data/HFP/wilderness/stressors"), full = T)

for(i in 2:length(years)){
  print(years[[i]])
  
  stress<-c(wild_stress[grep(pattern = years[[i]], x = wild_stress)], wild_chi[grep(pattern = years[[i]], wild_chi)])
  
  chi_stack<-rast(c(stress))

  chi_sum<-app(chi_stack, sum)

  chi_final<-chi_sum
  chi_final[chi_final>0]<-1 # changing not wilderness to 1

  writeRaster(chi_final, paste0(here("output_data/HFP/wilderness/direct_wilderness/Direct_high_impact_10perc_"), years[[i]],".tif"))
  
}
  
  


```




Values of 1 above indicate areas that are not wilderness, values of 0 indicate wilderness.


The final layers will then be the stoppable human threat across years.

```{r}
test<-rast(here("output_data/HFP/wilderness/direct_wilderness/Direct_high_impact_10perc_2013.tif"))
```



