---
title: "Human_footprint"
author: "Jade"
date: "2024-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load packages
```{r}
library(sf)
library(here)
library(tidyverse)
library(lubridate)
library(terra)
```

All human footprint data downloaded from: https://knb.ecoinformatics.org/view/doi:10.5063/F12B8WBS


# Load data

```{r}
chi_2013 <- rast(here("raw_data/Human_footprint/Direct_high_impact_10perc_2013.tif"))
```

```{r}
plot(chi_2013)
```


```{r}
chi_2003 <- rast(here("raw_data/Human_footprint/Direct_high_impact_10perc_2003.tif"))
```

```{r}
plot(chi_2003)
```

#Find threshold value for converted waters 
Ocean wilderness has previously been considered the 10% of ocean with least human impacts - we could find the least impacted 10% of aus waters 

#change raster to 0 and 1 
This code is for an example, reclassing numbers less than the mean as 1 and more than the mean as 2. Needs to be as a matrix 
reclass_matrix <- c(0, sst_mean, 1, 
                    sst_mean, Inf, 2)%>% 
                    matrix(ncol = 3, byrow=TRUE)
                    
reclassed <- classify(raster, reclass_matrix)
                    