---
title: "Human_footprint"
author: "Jade"
date: "2024-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load packages
```{r}
library(sf)
library(here)
library(tidyverse)
library(lubridate)
library(terra)
```

# Load data
Downloaded from: https://knb.ecoinformatics.org/view/doi:10.5063/F12B8WBS
2003 
```{r}
footprint2003 <- rast(here("raw_data/Human_footprint/cumulative_impact_2003.tif"))
```

```{r}
plot(log10(footprint2003+1))
```

#Load in data set 2013
```{r}
footprint2013 <- rast(here("raw_data/Human_footprint/cumulative_impact_2013.tif"))
```


#Reproject human footprint data 
```{r}
footprint2013_projected<- project(footprint2013,"epsg:3577")
```

#Save projection 
```{r}
writeRaster(footprint2013_projected, here("output_data/2013_footprint_projected.tif"), overwrite = FALSE)
```

#Load in EEZ data 
```{r}
eez<- vect(here("raw_data/EEZ/au_eez_pol_april2022.shp"))
```

##Reproject EEZ to epsg:3577 
```{r}
eez_project <- project(eez, "epsg:3577")
```

#Input capad data 
```{r}
capad<- vect(here("output_data/All_MPA_baseline.shp"))

strict<- vect(here("output_data/Strict_baseline.shp"))

non_strict<- vect(here("output_data/Non_strict_baseline.shp"))

```

##Plot to check projections line up 
```{r}
plot(log10(footprint2013_projected+1))
lines(capad)
lines(eez_project)

#This looks strange due to Aus projection, but the capad + footprint data sets line up  
```


#Clip footprint to Aus EEZ
```{r}
aus_footprint_2013<- crop(footprint2013_projected, eez_project, mask = TRUE)

```

##Plot aus data 
```{r}
plot(aus_footprint_2013)
lines(capad)
```


#Use zonal statistics to find average impact in strict vs non-strict vs open 
Can use hist() or freq() to see the break down of raster values 
freq(, value = NA) to show how many are NA

##Average for aus EEZ
```{r}
global(aus_footprint_2013, "mean", na.rm = TRUE)

#Mean = 1.243893
```


global() gives statistics e.g. global(raster, "mean", na.rm = TRUE)
##footprint for all protected areas
```{r}
capad_footprint_2013<- crop(aus_footprint_2013, capad, mask= TRUE)

global(capad_footprint_2013, "mean", na.rm = TRUE)
#mean 1.164847

hist(capad_footprint_2013)
```

## footprint for strict PA
```{r}
strict_footprint_2013 <- crop(aus_footprint_2013, strict, mask = TRUE)

global(strict_footprint_2013, "mean", na.rm = TRUE)
#Mean 1.137427 
```

## footprint for non strict PA 
```{r}
nonstrict_footprint_2013 <- crop(aus_footprint_2013, non_strict, mask = TRUE)

global(nonstrict_footprint_2013, "mean", na.rm = TRUE)
#Mean 1.180736
```


#Select out open ocean
```{r}
open_footprint_2013 <- mask(aus_footprint_2013, capad, inverse = TRUE)

#plot(open_footprint_2013)
#lines(capad)

```

## footprint for open areas
```{r}
global(open_footprint_2013, "mean", na.rm = TRUE)
#mean = 1.313475

```


#Find threshold value for converted waters 
Ocean wilderness has previously been considered the 10% of ocean with least human impacts - we could find the least impacted 10% of aus waters 

#change raster to 0 and 1 
This code is for an example, reclassing numbers less than the mean as 1 and more than the mean as 2. Needs to be as a matrix 
reclass_matrix <- c(0, sst_mean, 1, 
                    sst_mean, Inf, 2)%>% 
                    matrix(ncol = 3, byrow=TRUE)
                    
reclassed <- classify(raster, reclass_matrix)
                    