---
title: "Protection_Equality_Gap"
author: "Caitie"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(sf)
library(tidyverse)
library(here)
library(terra)
```


# IMCRA ecoregions

Downloaded from: https://fed.dcceew.gov.au/datasets/7394d7984e3f4e60990d4c455126bbfb/explore?layer=0

```{r}
imcra<-st_read(here("raw_data/IMCRA_regions/Integrated_Marine_and_Coastal_Regionalisation_of_Australia_(IMCRA)_v4.0_-_Meso-scale_Bioregions.shp"))
```


```{r}
imcra_diss<-imcra %>% 
  st_make_valid(.) %>% 
  st_transform(., "EPSG:3577") %>% 
  group_by(MESO_NAME) %>% 
  summarise(GIS_AREA_KM2 = sum(AREA_KM2, na.rm = T)) %>% 
  mutate(meso_area_km2 = as.numeric(st_area(.))*1e-6)

imcra_all <- imcra_diss %>%
mutate(ecoprotec = 0, prop_all = 0)
```

# CAPAD baseline

All

```{r}
capad<-st_read(here("output_data/Capad_baseline/All_MPA_baseline.shp"))
```


Strict

```{r}
strict<-st_read(here("output_data/Capad_baseline/Strict_baseline.shp")) %>% 
    filter(!is.na(Year) == TRUE) #perhaps think of another way to deal with missing dates

```


Non-strict

```{r}
non_strict<-st_read(here("output_data/Capad_baseline/Non_strict_baseline.shp"))
```


# Intersect ecoregions with MPA

```{r}

capad_files <- list.files(here("output_data/Capad_baseline"), pattern = "capad_baseline_", full.names = TRUE)

capad_files2<- capad_files[grep("shp", capad_files)]

years<- c(1916:2022)

for(i in 1:length(years)){
  print(years[[i]])
  sub_capad<- st_read(capad_files2[grep(years[[i]],capad_files2)])
  
  capad<-sub_capad %>% 
  mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)
  
  capad$new_capad_area<- capad%>% st_area(.)*1e-6
  
  baseline_capad_imcra<-st_intersection(capad, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_all = ecoprotec/meso_area_km2)  %>% 
     filter(st_geometry_type(.) !="LINESTRING")%>%
    filter(st_geometry_type(.) !="MULTILINESTRING")%>%
  st_cast(., "POLYGON")
  

meso_names_remove <- unique(baseline_capad_imcra$MESO_NAME)
filter_imcra_diss<- imcra_diss%>%
  filter(!MESO_NAME %in% meso_names_remove)%>%
  mutate( ecoprotec = 0,
    prop_all = 0)
  
final_baseline_capad_imcra <- bind_rows(baseline_capad_imcra, filter_imcra_diss)  
  
  st_write(final_baseline_capad_imcra, here(paste0("output_data/IMCRA_intersect/capad_imcra_baseline_",years[[i]],".shp")), append = FALSE)
}

```

```{r}
strict_files <- list.files(here("output_data/Capad_baseline"), pattern = "Strict_baseline", full.names = TRUE)

strict_only <- strict_files[!grepl("non",strict_files)]
  
strict_files2<- strict_only[grep("shp", strict_only)]

years<- c(1955:2022)
for(i in 1:length(years)){
  print(years[[i]])
  sub_strict<- st_read(strict_files2[grep(years[[i]],strict_files2)])
  
    strict<-sub_strict %>% 
  mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)
  
  baseline_strict_imcra<-st_intersection(strict, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_strict = ecoprotec/meso_area_km2)  %>% 
  st_cast(., "POLYGON")
  
  meso_names_remove <- unique(baseline_strict_imcra$MESO_NAME)
filter_imcra_diss<- imcra_diss%>%
  filter(!MESO_NAME %in% meso_names_remove)%>%
  mutate( ecoprotec = 0,
    prop_strict = 0)
  
final_baseline_strict_imcra <- bind_rows(baseline_strict_imcra, filter_imcra_diss)  

  st_write(final_baseline_strict_imcra, here(paste0("output_data/IMCRA_intersect/strict_imcra_baseline_",years[[i]],".shp")), append = TRUE)
}

```


```{r}
non_strict_files <- list.files(here("output_data/Capad_baseline"), pattern = "Non_strict_", full.names = TRUE)
  
non_strict_files2<- non_strict_files[grep("shp", non_strict_files)]

years<- c(1916:2022)

for(i in 2:length(years)){
  
  print(years[[i]])
  sub_non_strict<- st_read(non_strict_files2[grep(years[[i]],non_strict_files2)])
  
  non_strict<-sub_non_strict %>% 
  mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)
  
  baseline_non_strict_imcra<-st_intersection(non_strict, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
   filter(st_geometry_type(.) !="LINESTRING")%>%
    filter(st_geometry_type(.) !="MULTILINESTRING")%>%
  st_cast(.,"POLYGON")
  
  meso_names_remove <- unique(baseline_non_strict_imcra$MESO_NAME)
filter_imcra_diss<- imcra_diss%>%
  filter(!MESO_NAME %in% meso_names_remove)%>%
  mutate( ecoprotec = 0,
    prop_non_strict = 0)
  
final_baseline_non_strict_imcra <- bind_rows(baseline_non_strict_imcra, filter_imcra_diss)  

  
  st_write(final_baseline_non_strict_imcra, here(paste0("output_data/IMCRA_intersect/non_strict_imcra_baseline_",years[[i]],".shp")), append = FALSE)
}

```


```{r}
strict_imcra<-st_intersection(strict, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_strict = ecoprotec/meso_area_km2) %>% 
  st_cast(., "POLYGON")

non_strict_imcra<-st_intersection(non_strict, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")
```


# HERE YOU WOULD THEN REPEAT FOR THE CORRECTED ONE ONCE YOU HAVE DONE THE HUMAN FOOTPRINT CORRECTIONS

# CAPAD corrected
##Calculate ecoregion protected 

```{r}
capad_files <- list.files(here("output_data/Capad_adjusted"), pattern = "capad_adjust_", full.names = TRUE)

capad_files2<- capad_files[grep("shp", capad_files)]

years<- c(2003:2022)

for(i in 1:length(years)){
  print(years[[i]])
  sub_capad<- st_read(capad_files2[grep(years[[i]],capad_files2)])
  
  capad<-sub_capad %>% 
  mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)
  
  capad$new_capad_area<- capad%>% st_area(.)*1e-6
  
  adjust_capad_imcra<-st_intersection(capad, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_all = ecoprotec/meso_area_km2)  %>% 
     filter(st_geometry_type(.) !="LINESTRING")%>%
    filter(st_geometry_type(.) !="MULTILINESTRING")%>%
  st_cast(., "POLYGON")
  

meso_names_remove <- unique(adjust_capad_imcra$MESO_NAME)
filter_imcra_diss<- imcra_diss%>%
  filter(!MESO_NAME %in% meso_names_remove)%>%
  mutate( ecoprotec = 0,
    prop_all = 0)
  
final_adjust_capad_imcra <- bind_rows(adjust_capad_imcra, filter_imcra_diss)  
  
  st_write(final_adjust_capad_imcra, here(paste0("output_data/IMCRA_intersect_adjust/capad_imcra_adjusted_",years[[i]],".shp")), append = FALSE)
}

```


```{r}
strict_files <- list.files(here("output_data/Capad_adjusted"), pattern = "strict_adjust_2", full.names = TRUE)

strict_only <- strict_files[!grepl("non",strict_files)]
  
strict_files2<- strict_only[grep("shp", strict_only)]

years<- c(2003:2022)
for(i in 1:length(years)){
  print(years[[i]])
  sub_strict<- st_read(strict_files2[grep(years[[i]],strict_files2)])
  
    strict<-sub_strict %>% 
  mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)
  
  adjust_strict_imcra<-st_intersection(strict, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_strict = ecoprotec/meso_area_km2)  %>% 
  st_cast(., "POLYGON")
  
  meso_names_remove <- unique(adjust_strict_imcra$MESO_NAME)
filter_imcra_diss<- imcra_diss%>%
  filter(!MESO_NAME %in% meso_names_remove)%>%
  mutate( ecoprotec = 0,
    prop_strict = 0)
  
final_adjust_strict_imcra <- bind_rows(adjust_strict_imcra, filter_imcra_diss)  

  st_write(final_adjust_strict_imcra, here(paste0("output_data/IMCRA_intersect_adjust/strict_imcra_adjusted_",years[[i]],".shp")), append = FALSE)
}

```


```{r}
non_strict_files <- list.files(here("output_data/Capad_adjusted"), pattern = "non_strict_adjust_2", full.names = TRUE)
  
non_strict_files2<- non_strict_files[grep("shp", non_strict_files)]

years<- c(2003:2022)
for(i in 1:length(years)){
  print(years[[i]])
  sub_non_strict<- st_read(non_strict_files2[grep(years[[i]],non_strict_files2)])
  
  non_strict<-sub_non_strict %>% 
  mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)
  
  adjust_non_strict_imcra<-st_intersection(non_strict, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
   filter(st_geometry_type(.) !="LINESTRING")%>%
    filter(st_geometry_type(.) !="MULTILINESTRING")%>%
  st_cast(.,"POLYGON")
  
  meso_names_remove <- unique(adjust_non_strict_imcra$MESO_NAME)
filter_imcra_diss<- imcra_diss%>%
  filter(!MESO_NAME %in% meso_names_remove)%>%
  mutate( ecoprotec = 0,
    prop_non_strict = 0)
  
final_adjust_non_strict_imcra <- bind_rows(adjust_non_strict_imcra, filter_imcra_diss)  

  
  st_write(final_adjust_non_strict_imcra, here(paste0("output_data/IMCRA_intersect_adjust/non_strict_imcra_adjusted_",years[[i]],".shp")), append = FALSE)
}

```


#Summarise to find total percentage of ecoregion protected 
Baseline 

```{r}
strict_2022<- st_read(here("output_data/Capad_baseline/Full_strict_2022.shp"))

non_strict_2022<- st_read(here("output_data/Capad_baseline/Full_non_strict_2022.shp"))

capad_2022<- st_read(here("output_data/Capad_baseline/Full_capad_2022.shp"))


```

```{r}

strict_imcra_2022<-st_intersection(strict_2022, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_strict = ecoprotec/meso_area_km2) %>% 
  st_cast(., "POLYGON")

final_baseline_strict = strict_imcra_2022%>%
  select(MESO_NAME, meso_area_km2, ecoprotec, prop_strict)%>% 
  st_drop_geometry(.)


non_strict_imcra_2022<-st_intersection(non_strict_2022, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")

final_baseline_non_strict = non_strict_imcra_2022%>%
  select(MESO_NAME, meso_area_km2, ecoprotec, prop_non_strict)%>% 
  st_drop_geometry(.)


capad_imcra_2022<-st_intersection(capad_2022, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_capad = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")
final_baseline_capad = capad_imcra_2022%>%
  select(MESO_NAME, meso_area_km2, ecoprotec, prop_capad)%>% 
  st_drop_geometry(.)

st_write(strict_imcra_2022, here("output_data/results/Strict_baseline_imcra_2022.shp"))
write.csv(strict_imcra_2022, here("output_data/results/Strict_baseline_imcra_2022.csv"))
write.csv(final_baseline_strict, here("output_data/results/Final_baseline_strict_imcra.csv"))

st_write(non_strict_imcra_2022, here("output_data/results/Non_strict_baseline_imcra_2022.shp"))
write.csv(non_strict_imcra_2022, here("output_data/results/Non_strict_baseline_imcra_2022.csv"))
write.csv(final_baseline_non_strict, here("output_data/results/Final_baseline_non_strict_imcra.csv"))


st_write(capad_imcra_2022, here("output_data/results/Capad_baseline_imcra_2022.shp"))
write.csv(capad_imcra_2022, here("output_data/results/Capad_baseline_imcra_2022.csv"))

write.csv(final_baseline_capad, here("output_data/results/Final_baseline_capad_imcra.csv"))


```


Corrected

```{r}
all_capad<- st_read(here("output_data/Capad_adjusted/capad_adjust_2022.shp"))

adjust_capad<- all_capad%>%
   mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)

all_strict<- st_read(here("output_data/Capad_adjusted/strict_adjust_2022.shp"))
adjust_strict<- all_strict%>%
   mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)

all_non_strict<- st_read(here("output_data/Capad_adjusted/non_strict_adjust_2022.shp"))
adjust_non_strict<- all_non_strict%>%
   mutate(diss = 1) %>% 
  group_by(diss) %>% 
  summarise(.)

```

```{r}

adjust_capad_imcra<-st_intersection(adjust_capad, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_capad = ecoprotec/meso_area_km2) %>% 
  st_cast(., "POLYGON")

final_adjust_capad = adjust_capad_imcra%>%
  select(MESO_NAME, meso_area_km2, ecoprotec, prop_capad)%>% 
  st_drop_geometry(.)


adjust_strict_imcra<-st_intersection(adjust_strict, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_strict = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")

final_adjust_strict = adjust_strict_imcra%>%
  select(MESO_NAME, meso_area_km2, ecoprotec, prop_strict)%>% 
  st_drop_geometry(.)


adjust_non_strict_imcra<-st_intersection(adjust_non_strict, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")

final_adjust_non_strict = adjust_non_strict_imcra%>%
  select(MESO_NAME, meso_area_km2, ecoprotec, prop_non_strict)%>% 
  st_drop_geometry(.)


```


```{r}
#st_write(adjust_strict_imcra, here("output_data/results/Strict_adjust_imcra.shp"))
write.csv(final_adjust_strict, here("output_data/results/Final_strict_adjust_imcra.csv"))

#st_write(adjust_non_strict_imcra, here("output_data/results/Non_strict_adjust_imcra.shp"))
write.csv(final_adjust_non_strict, here("output_data/results/Final_non_strict_adjust_imcra.csv"))

#st_write(adjust_capad_imcra, here("output_data/results/Capad_adjust_imcra.shp"))
write.csv(final_adjust_capad, here("output_data/results/Final_capad_adjust_imcra.csv"))

```


##No. of ecoregions with protection greater than 30% goal 
```{r}
count_strict_base<- nrow(subset(strict_imcra_2022, strict_imcra_2022$prop_strict>= 0.3))
#1 
count_non_strict_base<- nrow(subset(non_strict_imcra_2022, non_strict_imcra_2022$prop_non_strict>= 0.3))
#23

count_capad_base<- nrow(subset(capad_imcra_2022, capad_imcra_2022$prop_capad>= 0.3))
#28


count_strict_adjust<- nrow(subset(adjust_strict_imcra, adjust_strict_imcra$prop_strict>= 0.3))
#1 
count_non_strict_base<- nrow(subset(adjust_non_strict_imcra, adjust_non_strict_imcra$prop_non_strict>= 0.3))
#22

count_capad_base<- nrow(subset(adjust_capad_imcra, adjust_capad_imcra$prop_capad>= 0.3))
#27

```

