---
title: "Protection_Equality_Gap"
author: "Caitie"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(sf)
library(tidyverse)
library(here)
```


# IMCRA ecoregions

Downloaded from: https://fed.dcceew.gov.au/datasets/7394d7984e3f4e60990d4c455126bbfb/explore?layer=0

```{r}
imcra<-st_read(here("raw_data/IMCRA_regions/Integrated_Marine_and_Coastal_Regionalisation_of_Australia_(IMCRA)_v4.0_-_Meso-scale_Bioregions.shp"))
```


```{r}
imcra_diss<-imcra %>% 
  st_make_valid(.) %>% 
  st_transform(., "EPSG:3577") %>% 
  group_by(MESO_NAME) %>% 
  summarise(GIS_AREA_KM2 = sum(AREA_KM2, na.rm = T)) %>% 
  mutate(meso_area_km2 = as.numeric(st_area(.))*1e-6)
```

# CAPAD baseline

All

```{r}
capad<-st_read(here("output_data/Capad_baseline/All_MPA_baseline.shp"))
```


Strict

```{r}
strict<-st_read(here("output_data/Capad_baseline/Strict_baseline.shp")) %>% 
    filter(!is.na(Year) == TRUE) #perhaps think of another way to deal with missing dates

```


Non-strict

```{r}
non_strict<-st_read(here("output_data/Capad_baseline/Non_strict_baseline.shp"))
```


# Intersect ecoregions with MPA

```{r}
capad_imcra<-st_intersection(capad, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_all = ecoprotec/meso_area_km2) %>% #mutate is to make a new column. 
  #this finds the proportion area protected of each ecoregion 
  #the . in st_area allows us to use the output from the intersecting function.
  st_cast(., "POLYGON")
```


```{r}
strict_imcra<-st_intersection(strict, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_strict = ecoprotec/meso_area_km2) %>% 
  st_cast(., "POLYGON")

non_strict_imcra<-st_intersection(non_strict, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")
```

```{r}
st_write(capad_imcra, here("output_data/IMCRA_intersect/All_capad_imcra_intersect.shp"), append = FALSE)


st_write(strict_imcra, here("output_data/IMCRA_intersect/Strict_imcra_intersect.shp"), append = FALSE)

view(st_read(here("output_data/IMCRA_intersect/All_capad_imcra_intersect.shp")))


st_write(non_strict_imcra, here("output_data/IMCRA_intersect/Non_strict_imcra_intersect.shp"), append = FALSE)
```


# HERE YOU WOULD THEN REPEAT FOR THE CORRECTED ONE ONCE YOU HAVE DONE THE HUMAN FOOTPRINT CORRECTIONS

# CAPAD corrected
## combine into one data set with each year 
```{r}
base<-st_read(here("output_data/Capad_adjusted/capad_adjust_2003.shp"))

capad_files <- list.files(here("output_data/Capad_adjusted"), pattern = "capad_adjust", full.names = TRUE)

capad_files2<- capad_files[grep("shp", capad_files)]

years<- c(2004:2013)

for(i in 1:length(years)){
  print(years[[i]])
  sub_capad<- st_read(capad_files2[grep(years[[i]],capad_files2)])

  select_year <- sub_capad[sub_capad$Year == years[[i]],]
  updated_capad <- rbind(base,select_year)
  base <- updated_capad
}

# Save the combined shapefile
st_write(base, here("output_data/Capad_adjusted/All_capad_adjusted.shp"), append = FALSE)
```


```{r}
base_strict<-st_read(here("output_data/Capad_adjusted/strict_adjust_2003.shp"))

strict_files <- list.files(here("output_data/Capad_adjusted"), pattern = "strict_adjust", full.names = TRUE)

strict_only <- strict_files[!grepl("non",strict_files)]
  
strict_files2<- strict_only[grep("shp", strict_only)]


years<- c(2004:2013)

for(i in 1:length(years)){
  print(years[[i]])
  sub_strict<- st_read(strict_files2[grep(years[[i]],strict_files2)])

  select_year <- sub_strict[sub_strict$Year == years[[i]],]
  updated_strict <- rbind(base_strict,select_year)
  base_strict <- updated_strict
}

# Save the combined shapefile
st_write(base_strict, here("output_data/Capad_adjusted/All_strict_adjusted.shp"), append = FALSE)
```

```{r}
base_non<-st_read(here("output_data/Capad_adjusted/non_strict_adjust_2003.shp"))

non_strict_files <- list.files(here("output_data/Capad_adjusted"), pattern = "non_strict_adjust", full.names = TRUE)
  
non_strict_files2<- non_strict_files[grep("shp", non_strict_files)]


non_years<- c(2004:2013)

for(i in 1:length(years)){
  print(years[[i]])
  sub_non<- st_read(non_strict_files2[grep(years[[i]],non_strict_files2)])

  select_year <- sub_non[sub_non$Year == years[[i]],]
  updated_non <- rbind(base_non,select_year)
  base_non <- updated_non
}

# Save the combined shapefile
st_write(base_non, here("output_data/Capad_adjusted/All_non_strict_adjusted.shp"), append = FALSE)
```

```{r}
capad_adjust<- st_read(here("output_data/Capad_adjusted/All_capad_adjusted.shp"))
strict_adjust<- st_read(here("output_data/Capad_adjusted/All_strict_adjusted.shp"))
non_strict_adjust<- st_read(here("output_data/Capad_adjusted/All_non_strict_adjusted.shp"))
```

##Calculate ecoregion protected 

```{r}
capad_files <- list.files(here("output_data/Capad_adjusted"), pattern = "capad_adjust_", full.names = TRUE)

capad_files2<- capad_files[grep("shp", capad_files)]

years<- c(2003:2013)
for(i in 1:length(years)){
  print(years[[i]])
  sub_capad<- st_read(capad_files2[grep(years[[i]],capad_files2)])
  
  adjust_capad_imcra<-st_intersection(sub_capad, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_all = ecoprotec/meso_area_km2)  %>% 
     filter(st_geometry_type(.) !="LINESTRING")%>%
    filter(st_geometry_type(.) !="MULTILINESTRING")%>%
  st_cast(., "POLYGON")
  
  st_write(adjust_capad_imcra, here(paste0("output_data/IMCRA_intersect_adjust/capad_imcra_adjusted_",years[[i]],".shp")), append = FALSE)
}

```


```{r}
strict_files <- list.files(here("output_data/Capad_adjusted"), pattern = "strict_adjust", full.names = TRUE)

strict_only <- strict_files[!grepl("non",strict_files)]
  
strict_files2<- strict_only[grep("shp", strict_only)]

years<- c(2003:2013)
i =1 
for(i in 1:length(years)){
  print(years[[i]])
  sub_strict<- st_read(strict_files2[grep(years[[i]],strict_files2)])
  
  adjust_strict_imcra<-st_intersection(sub_strict, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_strict = ecoprotec/meso_area_km2)  %>% 
  st_cast(., "POLYGON")
  
  st_write(adjust_strict_imcra, here(paste0("output_data/IMCRA_intersect_adjust/strict_imcra_adjusted_",years[[i]],".shp")), append = FALSE)
}

```


```{r}
non_strict_files <- list.files(here("output_data/Capad_adjusted"), pattern = "non_strict_adjust", full.names = TRUE)
  
non_strict_files2<- non_strict_files[grep("shp", non_strict_files)]

years<- c(2003:2013)
for(i in 1:length(years)){
  print(years[[i]])
  sub_non_strict<- st_read(non_strict_files2[grep(years[[i]],non_strict_files2)])
  
  adjust_non_strict_imcra<-st_intersection(sub_non_strict, imcra_diss) %>% 
  st_make_valid(.) %>%
    mutate(ecoprotec = as.numeric(st_area(.))*1e-6,
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
   filter(st_geometry_type(.) !="LINESTRING")%>%
    filter(st_geometry_type(.) !="MULTILINESTRING")%>%
  st_cast(.,"POLYGON")
  
  st_write(adjust_non_strict_imcra, here(paste0("output_data/IMCRA_intersect_adjust/non_strict_imcra_adjusted_",years[[i]],".shp")), append = FALSE)
}

```


#Summarise to find total percentage of ecoregion protected 
Baseline 

```{r}
strict_2022<- st_read(here("output_data/Capad_baseline/Full_strict_2022.shp"))

non_strict_2022<- st_read(here("output_data/Capad_baseline/Full_non_strict_2022.shp"))

```

```{r}

strict_imcra_2022<-st_intersection(strict_2022, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_strict = ecoprotec/meso_area_km2) %>% 
  st_cast(., "POLYGON")

non_strict_imcra_2022<-st_intersection(non_strict_2022, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")

st_write(strict_imcra_2022, here("output_data/results/Strict_fullnetwork_imcra_2022.shp"))

st_write(non_strict_imcra_2022, here("output_data/results/Non_strict_fullnetwork_imcra_2022.shp"))


```


Corrected

```{r}
correct_strict_2013<- st_read(here("output_data/Capad_adjusted/Full_strict_2013_corrected.shp"))

correct_non_strict_2013<- st_read(here("output_data/Capad_adjusted/Full_non_strict_2013_corrected.shp"))

```

```{r}

correct_strict_imcra_2013<-st_intersection(correct_strict_2013, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_strict = ecoprotec/meso_area_km2) %>% 
  st_cast(., "POLYGON")

correct_non_strict_imcra_2013<-st_intersection(correct_non_strict_2013, imcra_diss) %>% 
  st_make_valid(.) %>% 
  mutate(ecoprotec = as.numeric(st_area(.))*1e-6, 
         prop_non_strict = ecoprotec/meso_area_km2)%>% 
  st_cast(., "POLYGON")

```
##No. of ecoregions with protection greater than 30% goal 
```{r}
count_strict_base<- nrow(subset(strict_imcra_2022, strict_imcra_2022$prop_strict>= 0.3))
#1 
count_non_strict_base<- nrow(subset(non_strict_imcra_2022, non_strict_imcra_2022$prop_non_strict>= 0.3))
#23
```

