---
title: "Protection_Equality_Gap"
author: "Caitie"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(sf)
devtools::install_github("AChauvenet/ProtectEqual")
library(ProtectEqual)
library(tidyverse)
library(here)
```


#CAPAD IMCRA Intersect

All CAPAD MPA baseline

```{r}
capad_imcra<-st_read(here("output_data/CAPAD/All_capad_imcra_intersect.shp")) %>% 
  filter(!is.na(Year) == TRUE)
```


Strict baseline

```{r}
strict_imcra<-st_read(here("output_data/CAPAD/Strict_imcra_intersect.shp")) 
```


Non-strict baseline

```{r}
non_strict_imcra<-st_read(here("output_data/CAPAD/Non_strict_imcra_intersect.shp")) 
```


# Protection Equality calculation

Use the ProtectEqual package: https://github.com/AChauvenet/ProtectEqual
Need to cite it as: Chauvenet ALM, Kuempel CD and Possingham HP (2017). ProtectEqual. version 1.0.0. doi: 10.5281/zenodo.163951.

All capad baseline

```{r}
years<-min(capad_imcra$Year):max(capad_imcra$Year)

capad_pe<-data.frame(Year = NA, CAPAD_PE = NA, CAPAD_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-capad_imcra %>% filter(Year <= y)
  #y<-max(unique(sub$Year))
  print(y)
  sub2<-sub %>% 
    select(ms_r_k2, prop_ll) %>% 
    rename(ai = ms_r_k2,
           pi = prop_ll)
  pe<-pe(sub2, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  capad_pe[i,"Year"]<-y
  capad_pe[i, "CAPAD_PE"]<-pe$PE
  capad_pe[i, "CAPAD_N"]<-pe$N
}
```


Strict baseline
```{r}
strict_pe<-data.frame(Year = NA, Strict_PE = NA, Strict_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-strict_imcra %>% filter(Year <= y)
  print(y)
  sub2<-sub %>% 
    select(ms_r_k2, prp_str) %>% 
    rename(ai = ms_r_k2,
           pi = prp_str)
  pe<-pe(sub2, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  strict_pe[i,"Year"]<-y
  strict_pe[i, "Strict_PE"]<-pe$PE
  strict_pe[i, "Strict_N"]<-pe$N
}
```

Non-strict baseline

```{r}
non_strict_pe<-data.frame(Year = NA, Non_strict_PE = NA, Non_strict_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-non_strict_imcra %>% filter(Year <= y)
  print(y)
  sub2<-sub %>% 
    select(ms_r_k2, prp_nn_) %>% 
    rename(ai = ms_r_k2,
           pi = prp_nn_)
  pe<-pe(sub2, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  non_strict_pe[i,"Year"]<-y
  non_strict_pe[i, "Non_strict_PE"]<-pe$PE
  non_strict_pe[i, "Non_strict_N"]<-pe$N
}
```

```{r}
combine<-full_join(capad_pe, strict_pe, by = "Year") %>% 
  full_join(., non_strict_pe, by = "Year") %>% 
  arrange(., Year)
```

```{r}
write_csv(combine, here("output_data/results/PE_baseline_calculations.csv"))
```

# Protection Gap calculation

Strict baseline

Non-strict baseline
