---
title: "Protection_Equality_Gap"
author: "Caitie"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(sf)
#devtools::install_github("AChauvenet/ProtectEqual")
library(ProtectEqual)
library(dplyr)
library(tidyverse)
library(here)
#devtools::install_github("KerstinJantke/ConsTarget")
library(ConsTarget)
```

#CAPAD IMCRA Intersect

All CAPAD MPA baseline

```{r}
capad_imcra<-st_read(here("output_data/IMCRA_intersect/All_capad_imcra_intersect.shp")) %>% 
  filter(!is.na(Year) == TRUE)
```


Strict baseline

```{r}
strict_imcra<-st_read(here("output_data/IMCRA_intersect/Strict_imcra_intersect.shp")) 
```


Non-strict baseline

```{r}
non_strict_imcra<-st_read(here("output_data/IMCRA_intersect/Non_strict_imcra_intersect.shp")) 
```


# Protection Equality calculation

Use the ProtectEqual package: https://github.com/AChauvenet/ProtectEqual
Need to cite it as: Chauvenet ALM, Kuempel CD and Possingham HP (2017). ProtectEqual. version 1.0.0. doi: 10.5281/zenodo.163951.

All capad baseline

```{r}
years<-min(capad_imcra$Year):max(capad_imcra$Year)

capad_pe<-data.frame(Year = NA, CAPAD_PE = NA, CAPAD_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-capad_imcra %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T))
  
  print(y)
  pe<-pe(eco_sum, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  capad_pe[i,"Year"]<-y
  capad_pe[i, "CAPAD_PE"]<-pe$PE
  capad_pe[i, "CAPAD_N"]<-pe$N
}
```


Strict baseline
```{r}
strict_pe<-data.frame(Year = NA, Strict_PE = NA, Strict_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-strict_imcra %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T))
  pe<-pe(eco_sum, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  strict_pe[i,"Year"]<-y
  strict_pe[i, "Strict_PE"]<-pe$PE
  strict_pe[i, "Strict_N"]<-pe$N
}
```


Non-strict baseline

```{r}
non_strict_pe<-data.frame(Year = NA, Non_strict_PE = NA, Non_strict_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-non_strict_imcra %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
  
  eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T))
  
  pe<-pe(eco_sum, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  non_strict_pe[i,"Year"]<-y
  non_strict_pe[i, "Non_strict_PE"]<-pe$PE
  non_strict_pe[i, "Non_strict_N"]<-pe$N
}
```


```{r}
combine<-full_join(capad_pe, strict_pe, by = "Year") %>% 
  full_join(., non_strict_pe, by = "Year") %>% 
  arrange(., Year)
```


```{r}
write_csv(combine, here("output_data/results/PE_baseline_calculations.csv"))
```


# Protection Gap calculation
All capad areas 
```{r}
capad_mpg<-data.frame(Year = NA, Capad_MPG = NA, Capad_N = NA)

years<-min(capad_imcra$Year):max(capad_imcra$Year)


for(i in 1:length(years)){
   y<-years[[i]]
  sub<-capad_imcra %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T)) %>% 
  rename(feature = MESO_NA)

 mpg<- mpg(eco_sum, target = 0.3, plot = F)
  
  capad_mpg[i,"Year"]<-y
  capad_mpg[i, "Capad_MPG"]<-mpg$MPG
  capad_mpg[i, "Capad_N"]<-mpg$N
}
```



Strict baseline
Calculate the mean protection gap for 30%
```{r}

strict_mpg<-data.frame(Year = NA, Strict_MPG = NA, Strict_N = NA)

years<-min(strict_imcra$Year):max(strict_imcra$Year)


for(i in 1:length(years)){
   y<-years[[i]]
  sub<-strict_imcra %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T)) %>% 
  rename(feature = MESO_NA)

 mpg<- mpg(eco_sum, target = 0.3, plot = F)
  
  strict_mpg[i,"Year"]<-y
  strict_mpg[i, "Strict_MPG"]<-mpg$MPG
  strict_mpg[i, "Strict_N"]<-mpg$N
}
```


Non-strict baseline
```{r}
non_strict_mpg<-data.frame(Year = NA, Non_Strict_MPG = NA, Non_Strict_N = NA)

years<-min(non_strict_imcra$Year, na.rm = T):max(non_strict_imcra$Year, na.rm=T)

for(i in 1:length(years)){
   y<-years[[i]]
  sub<-non_strict_imcra %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T)) %>% 
    rename(feature = MESO_NA)

 mpg<- mpg(eco_sum, target = 0.1, plot = F)
  
  non_strict_mpg[i,"Year"]<-y
  non_strict_mpg[i, "Non_Strict_MPG"]<-mpg$MPG
  non_strict_mpg[i, "Non_Strict_N"]<-mpg$N
}
```

```{r}
combine_mpg<-full_join(capad_mpg, strict_mpg, by = "Year") %>% 
  full_join(., non_strict_mpg, by = "Year") %>% 
  arrange(., Year)
```


```{r}
write_csv(combine_mpg, here("output_data/results/MPG_baseline_calculations.csv"))
```


#Protection equality - corrected 

```{r}
capad_imcra_adjusted<- st_read(here("output_data/Adjust_all_capad_imcra_intersect.shp"))%>% 
  filter(!is.na(Year) == TRUE)


strict_imcra_adjusted<- st_read(here("output_data/Adjust_strict_imcra_intersect.shp"))%>% 
  filter(!is.na(Year) == TRUE)


non_strict_imcra_adjusted<- st_read( here("output_data/Adjust_non_strict_imcra_intersect.shp"))%>% 
  filter(!is.na(Year) == TRUE)
```



```{r}
years<-min(capad_imcra_adjusted$Year):max(capad_imcra_adjusted$Year)

adjust_capad_pe<-data.frame(Year = NA, CAPAD_PE = NA, CAPAD_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-capad_imcra_adjusted %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T))
  #y<-max(unique(sub$Year))
  print(y)
  pe<-pe(eco_sum, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  adjust_capad_pe[i,"Year"]<-y
  adjust_capad_pe[i, "CAPAD_PE"]<-pe$PE
  adjust_capad_pe[i, "CAPAD_N"]<-pe$N
}
```



Strict corrected
```{r}
adjust_strict_pe<-data.frame(Year = NA, Strict_PE = NA, Strict_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-strict_imcra_adjusted %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T))
  pe<-pe(eco_sum, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  adjust_strict_pe[i,"Year"]<-y
  adjust_strict_pe[i, "Strict_PE"]<-pe$PE
  adjust_strict_pe[i, "Strict_N"]<-pe$N
}
```


Non-strict corrected

```{r}
adjust_non_strict_pe<-data.frame(Year = NA, Non_strict_PE = NA, Non_strict_N = NA)

for(i in 1:length(years)){
  y<-years[[i]]
  sub<-non_strict_imcra_adjusted %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
  
  eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T))
  
  pe<-pe(eco_sum, version = "proportional", plot_pe = FALSE, correct = TRUE)
  
  adjust_non_strict_pe[i,"Year"]<-y
  adjust_non_strict_pe[i, "Non_strict_PE"]<-pe$PE
  adjust_non_strict_pe[i, "Non_strict_N"]<-pe$N
}
```


```{r}
combine_adjust<-full_join(adjust_capad_pe, adjust_strict_pe, by = "Year") %>% 
  full_join(., adjust_non_strict_pe, by = "Year") %>% 
  arrange(., Year)
```


```{r}
write_csv(combine_adjust, here("output_data/results/PE_corrected_calculations.csv"))
```


#MPG - corrected
```{r}
adjust_capad_mpg<-data.frame(Year = NA, Capad_MPG = NA, Capad_N = NA)

years<-min(capad_imcra_adjusted$Year):max(capad_imcra_adjusted$Year)


for(i in 1:length(years)){
   y<-years[[i]]
  sub<-capad_imcra_adjusted %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T)) %>% 
  rename(feature = MESO_NA)

 mpg<- mpg(eco_sum, target = 0.3, plot = F)
  
  adjust_capad_mpg[i,"Year"]<-y
  adjust_capad_mpg[i, "Capad_MPG"]<-mpg$MPG
  adjust_capad_mpg[i, "Capad_N"]<-mpg$N
}
```



Strict baseline
Calculate the mean protection gap for 30%
```{r}

adjust_strict_mpg<-data.frame(Year = NA, Strict_MPG = NA, Strict_N = NA)

years<-min(strict_imcra_adjusted$Year):max(strict_imcra_adjusted$Year)


for(i in 1:length(years)){
   y<-years[[i]]
  sub<-strict_imcra_adjusted %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T)) %>% 
  rename(feature = MESO_NA)

 mpg<- mpg(eco_sum, target = 0.3, plot = F)
  
  adjust_strict_mpg[i,"Year"]<-y
  adjust_strict_mpg[i, "Strict_MPG"]<-mpg$MPG
  adjust_strict_mpg[i, "Strict_N"]<-mpg$N
}
```


Non-strict baseline
```{r}
adjust_non_strict_mpg<-data.frame(Year = NA, Non_Strict_MPG = NA, Non_Strict_N = NA)

years<-min(non_strict_imcra_adjusted$Year, na.rm = T):max(non_strict_imcra_adjusted$Year, na.rm=T)

for(i in 1:length(years)){
   y<-years[[i]]
  sub<-non_strict_imcra_adjusted %>% filter(Year <= y) %>%
    as.data.frame() %>%
    dplyr::select(-geometry)
  print(y)
eco_sum<-sub %>%
    group_by(MESO_NA) %>%
    summarise(ai = unique(ms_r_k2),
              pi = sum(ecoprtc, na.rm = T)) %>% 
    rename(feature = MESO_NA)

 mpg<- mpg(eco_sum, target = 0.1, plot = F)
  
  adjust_non_strict_mpg[i,"Year"]<-y
 adjust_non_strict_mpg[i, "Non_Strict_MPG"]<-mpg$MPG
  adjust_non_strict_mpg[i, "Non_Strict_N"]<-mpg$N
}
```

```{r}
combine_mpg<-full_join(adjust_capad_mpg, adjust_strict_mpg, by = "Year") %>% 
  full_join(., adjust_non_strict_mpg, by = "Year") %>% 
  arrange(., Year)
```


```{r}
write_csv(combine_mpg, here("output_data/results/MPG_corrected_calculations.csv"))
```

